Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-30 inst/paper/main.tex
sec 2.3

\begin{lstlisting}
 eval
   <- function(expr,
                envir = parent.frame(),
                enclos = if(is.list(envir)) parent.frame() else baseenv())
          { ...body... }
\end{lstlisting}
The parameters are \c{expr}, the value to be evaluated, \c{envir}, the
environment in which evaluation happens, and \c{enclos}, the
environment to look up objects not found in \c{envir} (by default
either the top level, \c{baseenv} or the environment of the caller).
The \c{expr} parameter can take values of different type, for our
purposes we focus on the most common, {\tt expression}s. It is easiest
to think of an \c{expression} as an abstract syntax tree as returned
by R's parser. The simplest way to create an \c{expression} is to call
\c{quote} passing some R expression:
\begin{lstlisting}
  > quote(a + b)
  # a + b
\end{lstlisting}
Here the return value is an abstract syntax tree. From a
string, use  the \c{parse} function:
\begin{lstlisting}
  > parse(text="a+b")
  # expression(a+b)
\end{lstlisting}
The most common way to create an expression is to extract it from an
argument. Each argument comes packaged inside a promise which retains
the source code of the argument. Consider the following function
definition, \c{f} is a function with a single parameter \c{x}. When
the call \c{f(a+b)} is evaluated, \c{a+b} is stored inside a newly
created promise.
\begin{lstlisting}
  > f <- function(x) substitute(x), list(a=1)))
  > f(a+b)
  # expression(1+b)
\end{lstlisting}
The call to \c{substitute(x)} extracts, from the promise bound to
\c{x}, the expression corresponding to that argument, i.e., \c{a+b}.
The general form of the function takes two arguments, the second
argument is an object that can be used as an environment. This can be
either a list of named values, a data frame or an actual environment.
\begin{lstlisting}
  > environment()
  # <environment: R_GlobalEnv>
\end{lstlisting}
The above shows how to access the current environment, this can belong
to the current function or, as here, the top level. Environments are
nested, with each environment having a parent. This nesting is used
when looking up names. When calling \c{new.env}, the default parent is
the current environment. The chain of environments can be traversed
with \c{parent.env}, ultimately one arrives at \c{emptyenv}.
\c{parent.frame} and \c{sys.frame} grant access to environments
further up in the calling stack, up to the global environment
\c{.GlobalEnv}, and packages. One can also directly read, modify or
create new bindings, given any environment:
\begin{itemize}
\item \c{env$v} and \c{get("v", envir = env)} read variable \c{v} in
  \c{e};
\item \c{env$v <- 2} and \c{assign("v", 2, envir = env)} write 2 in
  \c{v}. If it did not exist, \c{v} is created in \c{env}.
\end{itemize}
\noindent Environments are often used as hashtables by programmers as
they have reference semantics (other are copy-on-write) and have a
built-in string look up. across function calls, or to create package
namespaces.
\begin{lstlisting}
  evalq <- function(exp,env,enclos) eval(quote(exp),env,enclos)
  eval.parent <- function(exp,n) eval(exp,parent.frame(n))
  local <- function(exp,env,enclos) evalq(exp,new.env())
\end{lstlisting}
The above code show simplified implementation for the variants of
\eval. The \c{evalq} form simply quotes the argument to prevent it
from being evaluated in the current environment. The
\c{eval.parent(e,n)} form evaluates \c{e} in the environment of the
  {\tt n}-th caller up the call stack. Finally \c{local} evaluates an
  expression in a new environment to avoid polluting the current
  environment.

%% Taking an expression rather
%% than a string, may discourage some casual users as manipulating
%% expressions is slightly more involved.\footnote{The rationale for
%% \eval taking expressions is explained by the main use case for \eval,
%% namely extending the language by reinterpretation of promises; R gives
%% access to the code of a promise as an expression.}

>>>>

\begin{lstlisting}
 eval
   <- function(expr,
                envir = parent.frame(),
                enclos = if(is.list(envir)) parent.frame() else baseenv())
          { ...body... }
\end{lstlisting}
The parameters are \c{expr}, the value to be evaluated, \c{envir}, the
environment in which evaluation happens, and \c{enclos}, the
environment to look up objects not found in \c{envir} (by default
either the top level, \c{baseenv} or the environment of the caller).
The \c{expr} parameter can take values of different type, for our
purposes we focus on the most common, {\tt expression}s. It is easiest
to think of an \c{expression} as an abstract syntax tree as returned
by R's parser. The simplest way to create an \c{expression} is to call
\c{quote} passing some R expression:
\begin{lstlisting}
  > quote(a + b)
  # a + b
\end{lstlisting}
Here the return value is an abstract syntax tree. From a
string, use  the \c{parse} function:
\begin{lstlisting}
  > parse(text="a+b")
  # expression(a+b)
\end{lstlisting}
The most common way to create an expression is to extract it from an
argument. Each argument comes packaged inside a promise which retains
the source code of the argument. Consider the following function
definition, \c{f} is a function with a single parameter \c{x}. When
the call \c{f(a+b)} is evaluated, \c{a+b} is stored inside a newly
created promise.
\begin{lstlisting}
  > f <- function(x) substitute(x), list(a=1)))
  > f(a+b)
  # expression(1+b)
\end{lstlisting}
The call to \c{substitute(x)} extracts, from the promise bound to
\c{x}, the expression corresponding to that argument, i.e., \c{a+b}.
The general form of the function takes two arguments, the second
argument is an object that can be used as an environment. This can be
either a list of named values, a data frame or an actual environment.
\begin{lstlisting}
  > environment()
  # <environment: R_GlobalEnv>
\end{lstlisting}
The above shows how to access the current environment, this can belong
to the current function or, as above, the top level. Environments are
nested, with each environment having a parent. This nesting is used
when looking up names. When calling \c{new.env}, the default parent is
the current environment. The chain of environments can be traversed
with \c{parent.env}, ultimately one arrives at \c{emptyenv}.
\c{parent.frame} and \c{sys.frame} grant access to environments
further up in the calling stack, up to the global environment
\c{.GlobalEnv}, and packages. One can also directly read, modify or
create new bindings, given any environment:
\begin{itemize}
\item \c{env$v} and \c{get("v",envir=env)} read variable \c{v} in
  environment \c{env};
\item \c{env$v<-2} and \c{assign("v",2,envir=env)} write 2 in
  \c{v}. If not found, \c{v} is created in \c{env}.
\end{itemize}
\noindent Environments are often used as hashtables by programmers as
they have reference semantics (other are copy-on-write) and have a
built-in string look up. across function calls, or to create package
namespaces.
\begin{lstlisting}
  evalq <- function(exp,env,enclos) eval(quote(exp),env,enclos)
  eval.parent <- function(exp,n) eval(exp,parent.frame(n))
  local <- function(exp,env,enclos) evalq(exp,new.env())
\end{lstlisting}
The above code show simplified implementation for the variants of
\eval. The \c{evalq} form simply quotes the argument to prevent it
from being evaluated in the current environment. The
\c{eval.parent(e,n)} form evaluates \c{e} in the environment of the
  {\tt n}-th caller up the call stack. Finally, \c{local} evaluates an
  expression in a new environment to avoid polluting the current
  environment.

\subsection{Eval Usage in R}

The R language was intended to be extensible, the combination of lazy
evaluation, \c{substitute} and \c{eval} are the tools given to
developers to this end. This API is slightly more complex than just
passing a string, it is conceivable that this may discourage some
casual users. \Eval is also being used to reduce boilerplate code and
provide convenience features for programmers. We now give some
representative examples of its usage.
